---
title: "US Airport Delay and snowfall in January 2018"
author: "Peter Jiyuan Zhou"
subtitle: A study to address airport delay and snow
---

# Introduction
Air travel is increasingly popular in not only the United States but also other parts of the world. However, flight delay are getting more and more normal especially for large airports. In the northern part of the US, snow is often one of the factor resulting in flight delay in winter and it is especially annoying. This project aims to analyse the vulnerability of airport delays to snowfall across different region in the US.  

# Data
* [On Time Performance Data](https://www.transtats.bts.gov/Tables.asp?DB_ID=120&DB_Name=Airline%20On-Time%20Performance%20Data&DB_Short_Name=On-Time): "Airline On Time Performance Data" from Bureau of Transportation Statistics, Unitd States Department of Transportation.
* [Airport Location Data](https://www.transtats.bts.gov/Tables.asp?DB_ID=595&DB_Name=Aviation%20Support%20Tables): "Master Coordinates" of "Aviation Support Table" from Bureau of Transportation Statistics, United States Department of Transportation

# Materials and methods

* Importing the data and data cleaning
```{r message = FALSE}
library(tidyverse)
library(ggmap)
library(ggplot2)
library(dplyr)
library(spData)
library(sf)
library(rgdal)
data(world)
US.sf <- world %>% dplyr::filter(iso_a2 == "US")
OTP <- read.csv("data/OnTimeP.csv")
APdata <- read.csv("data/AirportData.csv")
OTP.clean <- OTP %>% dplyr::filter(DEP_DELAY_NEW >= 1) %>% dplyr::filter(WEATHER_DELAY >= 1) %>% 
  dplyr::select(YEAR, MONTH, DAY_OF_MONTH, ORIGIN, ORIGIN_STATE_ABR, DEP_DELAY_NEW, WEATHER_DELAY)
OTP.sum <- OTP.clean %>% dplyr::select(ORIGIN, DEP_DELAY_NEW, WEATHER_DELAY) %>%
  group_by(ORIGIN) %>% summarize(total_delay = sum(DEP_DELAY_NEW), weather = sum(WEATHER_DELAY))
OTP.sum$ORIGIN <- sapply(OTP.sum$ORIGIN, as.character)
Airports <- APdata %>% select(AIRPORT, DISPLAY_AIRPORT_NAME, DISPLAY_AIRPORT_CITY_NAME_FULL, AIRPORT_COUNTRY_CODE_ISO, AIRPORT_STATE_CODE, LATITUDE, LONGITUDE)
A <- Airports %>% group_by(AIRPORT) %>% summarize(lat = mean(LATITUDE), lon = mean(LONGITUDE)) %>%
  dplyr::filter(lat != "NA") %>% dplyr::filter(lon != "NA")
A$AIRPORT <- sapply(A$AIRPORT, as.character)
Airport.sf <- st_as_sf(A, coords = c("lon", "lat"), crs = 4326)
```

# Results

* Visualization of Locaton of all US Airport considered
```{r}
A_O.inner <- left_join(OTP.sum, A, by = c("ORIGIN" = "AIRPORT"))
A_O.inner.sf <- st_as_sf(A_O.inner, coords = c("lon", "lat"), crs = 4326)
p1 <- ggplot(US.sf) + geom_sf(aes(geometry = geom)) + geom_point(data = A_O.inner, aes(x = lon, y = lat))
p1
```

* Visualization of US Airport with most weather related delays
```{r}
p2 <- ggplot(US.sf) + geom_sf(aes(geometry = geom)) + geom_point(data = A_O.inner, aes(x = lon, y = lat, size = weather))
p2
```

* Scattered plot of Total Delay in Minutes vs Weather Delay
```{r}
p3 <- ggplot(A_O.inner) + geom_point(aes(x = log(weather), y = log(total_delay))) + geom_smooth(method = loess, aes(x = log(weather), y = log(total_delay)))
p3
```


# Conclusions

What have you learned?  Are there any broader implications?

# References

